---
title: "HW1 final"
output:
  html_document:
    keep_md: yes
---

## Bias variance decomposition
```{r Bias variance decomposition}
library(ggplot2)
sq_bias = function(x) (2 * x - 20)^2 + 100
var = function(x)  exp(x / 2) * 4
train_error = function(x) (1.8*x - 18)^2 + 250
test_error = function(x) x^3/2 + 200 * cos(x/2) + 500
bayes_error = function(x) 200

line_width = 0.5
ggplot(data = data.frame(x = 0), mapping = aes(x = x)) + stat_function(fun = test_error, size = line_width, aes(color = "Test Error")) + stat_function(fun = sq_bias, size = line_width, aes(color = "Squared Bias")) + stat_function(fun = var, size = line_width, aes(color = "Variance")) +  stat_function(fun = train_error, size = line_width, aes(color = "Training Error")) +  stat_function(fun = bayes_error, size = line_width, aes(color = "Bayes Error")) + xlim(0, 8) + ylim(0, 800) + labs(title = "Typical Bias-Variance Decomposition", x="Flexibility(ascending)", y = "Error rate")
````
Generally as flexibility increases, bias decreases and variance increases. With higher flexibility, the model fits the observed data more closely, and the amount by which f would change if we estimated it using a different training set would increase (variance), but the error that is introduced by approximating real life problems using a simplified model decreases (bias). For training error, as flexibility increases, the model fits the training data better and training error decreases as a result. For test error, as flexibility increases, it has a u-shape, as it declines initially but increase again as the model has now overfit the data. For bayes or irreducible error, the error remains constant and whether a method is flexible or inflexible has no effect on it.

## Bayes classifier
```{r bayes classifier}
#creating the simulated dataset
set.seed(11) 
n = 200
x1 = runif(n, -1, 1)
x2 = runif(n, -1, 1)
d = data.frame(x1,x2)
epsilon = rnorm(n, mean = 0, sd = 0.5)
y = x1 + x1*x1 + x2 +x2*x2 + epsilon
y_new <- exp(y)/(1+exp(y))
y_new_d <- data.frame(x1, x2, y_new)
y_new_d$success = y_new > 0.5
y_new_d$success[y_new_d$success==TRUE] = "YES"
y_new_d$success[y_new_d$success==FALSE] = "NO"

##creating the grid/data for the plot
library(tidyverse)
bound = expand.grid(x1 = seq(-1, 1, by = 0.05), x2 = seq(-1, 1, by = 0.05))
bound = mutate(bound, ygrid = x1 + x1*x1 + x2 +x2*x2, pgrid = exp(ygrid)/(1+exp(ygrid)), grid_success = pgrid > 0.5)

#plotting the boundary
ggplot(bound, aes(x = x1, y = x2)) + geom_contour(aes(z = pgrid, group = grid_success), bins = 1) +
  geom_point(data = y_new_d, aes(color = success)) +
  scale_color_brewer(type = "qual", palette = 6) +
  labs(title = "Bayes decision boundary")
````
